Restrictive Clearance Comparison

### 2025-02-21 data-raw/RestrictiveClearance_comparison.Rmd
Included additional chemical descriptors and added relative log-likelihood calculations for restrictive/non-restrictive httk predictions.

```{r library_loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(ggplot2)
library(ctxR)
library(ggvenn)
library(ggalluvial)
library(flextable)
library(reticulate)
library(patchwork)
source("data-raw/RC_comparison_bkgd.R") # httk & dawson_2021 loaded here
```

@@

```{r venn_diagrams}
member_list <- list(
  httk_gas_pbtk = unique(get_cheminfo(info = "DTXSID",
                                species = "human",
                                model = "gas_pbtk",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)) %>% 
    suppressWarnings(),
  cvt = cvt %>% 
    filter(species %in% c("human", "rat"),
           conc_medium_normalized == "plasma") %>%
    pull(analyte_dtxsid)
)

@@

target_chems <- intersect(member_list$cvt, member_list$httk_gas_pbtk)

@@

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 12)
```

@@

@@

```{r lipinksi_test_data}
target_chem_info <- get_chemical_details_batch(DTXSID = target_chems,
                                               rate_limit = 1L)
target_chem_properties <- get_chem_info_batch(DTXSID = target_chems, 
                                              rate_limit = 1L) %>%
  filter(name == "LogKow: Octanol-Water") %>% distinct()

target_chem_properties %>%
  ggplot(aes(x = source, y = value)) +
  geom_point() +
  geom_line(aes(group = dtxsid), color = "grey10", alpha = 0.1) +
  geom_hline(yintercept = 5, linetype = "longdash") +
  facet_grid(cols = vars(propType), scales = "free_x") +
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, angle = 45))

@@

@@

```{python get_H_bonds}
import rdkit
from rdkit import Chem
from rdkit.Chem import Descriptors

@@

py_h_donors = [Chem.Descriptors.NOCount(x) for x in mol_from_smiles]
py_h_acceptors = [Chem.Descriptors.NHOHCount(x) for x in mol_from_smiles]
```

@@

target_chem_summary <- target_chem_info %>%
  distinct(dtxsid, monoisotopicMass, smiles) %>%
  right_join(data.frame(smiles = smiles_list)) %>%
  bind_cols(data.frame(H_donors = py_to_r(py$py_h_donors),
                       H_acceptors = py_to_r(py$py_h_acceptors))
            ) %>%
  left_join(
    target_chem_properties %>%
      filter(source == "ACD/Labs Consensus") %>%
      distinct(dtxsid, source, value)
  ) %>%
  rename(LogKow = value)

target_chem_lipinski <- target_chem_summary %>%
  select(-c(smiles, source)) %>%
  group_by(dtxsid) %>%
  mutate("Mass <= 500 Da" = (monoisotopicMass <= 500),
          "H-donors < 5" = (H_donors < 5),
          "H-acceptors < 10" = (H_acceptors < 10),
          "logP < 5" = (LogKow < 5),
          .keep = "none") %>%
  ungroup() %>%
  mutate(Lipinski = rowSums(dplyr::select(., where(is.logical))) >= 3)

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Alluvium_Lipinski.png"),
       bg = "white",
       height = 5, width = 6
)
```

@@

```{r httk_predictions}
restrictive_preds <- get_httk_preds(
  parameters = this_param_rest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_restrictive"),
  .cols = starts_with("httk_")
)

all_httk_preds <- inner_join(restrictive_preds, nonrestrictive_preds) %>%
  filter_targets()

@@

combined_preds <- all_httk_preds %>%
  left_join(winning_preds) %>%
  filter(!is.na(model))

combined_RMSLE <- combined_preds %>%
  ungroup() %>%
  mutate(
    restrictive_SLE = (log1p(httk_Preds_restrictive) - log1p(Conc))^2,
    nonrestrictive_SLE = (log1p(httk_Preds_nonrestrictive) - log1p(Conc))^2,
    invivoPK_SLE =  (log1p(Conc_est) - log1p(Conc))^2 
  ) %>%
  group_by(Chemical, Species) %>%
  summarize(RMSLE_restrictive = sqrt(mean(restrictive_SLE, na.rm = TRUE)),
            RMSLE_nonrestrictive = sqrt(mean(nonrestrictive_SLE, na.rm = TRUE)),
            RMSLE_invivoPKfit = sqrt(mean(invivoPK_SLE, na.rm = TRUE)))

@@

combined_RMSLE %>%
  ungroup() %>%
  mutate(comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
     between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive~ "More than"
  )) %>% 
  ggplot(aes(x = Chemical, y = RMSLE_invivoPKfit)) +
  geom_point(color = "green4") +
  geom_point(aes(y = RMSLE_nonrestrictive), shape = 6, size = 0.5) +
  geom_point(aes(y = RMSLE_restrictive), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = RMSLE_nonrestrictive, ymax = RMSLE_restrictive)) +
  facet_grid(~factor(comparison_group,
                     levels = c("Less than", "Between", "More than")),
             scales = "free_x", space = "free_x") +
  labs(y = "RMSLE") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "RMSLE_comparison_fullChart.png"),
       bg = "white",
       height = 4.5, width = 8
)

# Number of chemicals we have RMSLE's for
combined_RMSLE %>% pull(Chemical) %>% unique() %>% length()

@@

```{r relative-likelihood}
this_rll_probs <- AIC(this_pk) %>%
  filter_targets() %>%
  distinct(Chemical, Species, model, AIC) %>%
  pivot_wider(names_from = model, values_from = AIC) %>%
  rowwise() %>% 
  mutate(minAIC = min(c_across(starts_with("model_")), na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(across(starts_with("model_"), \(x) {
    signif(exp((minAIC - x)/2) * 100, 4)
  })) %>%
  mutate(across(starts_with("model_"), \(x) {
    ifelse(x < 1E-6, 0, x)
  })) %>%
  inner_join(this_nonflat_winmodel %>% distinct(Chemical, Species, model)) %>%
  ungroup() %>%
  filter(model_1comp >= 99 | model_2comp >= 99,
        model_flat < 1, Chemical %in% target_chems)

@@

@@

@@

```{r combine_tkstats}
# Get the winning model and only include chemicals in target_chems
this_winmodel <- get_winning_model(this_pk) %>%
    filter_targets()
# How many overlapping chemicals have winning models that aren't flat?
count(
  this_winmodel,
  model
) # 153 = 97 1-compartment + 56 2-compartment

this_tkstats <- eval_tkstats(this_pk,
                             model = c("model_1comp", "model_2comp"),
                             finite_only = TRUE,
                             dose_norm = TRUE
) %>%
  filter_targets() %>%
  rename(CLtot_Fgutabs.tkstats = `CLtot/Fgutabs.tkstats`) %>%
  distinct(Chemical, Species, method, model, Dose, Route, Media,
           CLtot.tkstats, CLtot_Fgutabs.tkstats, halflife.tkstats, Css.tkstats) %>%
  inner_join(this_winmodel)

@@

this_css_df <- inner_join(
  css_httk_batch(this_param_rest_l, restrictive = TRUE),
  css_httk_batch(this_param_nrest_l, restrictive = FALSE)
)
# Chart with relative differences (think steps) of Css between clearance types
this_css_df %>%
  mutate(relative_difference = log1p(Css_httk_r) - log1p(Css_httk_nr)) %>%
  count(relative_difference == 0, Css_httk_r == 0) # 31 Chemicals have zero difference
###
this_css_df %>%
  mutate(relative_difference = log1p(Css_httk_r) - log1p(Css_httk_nr)) %>%
  filter(relative_difference > 0) %>%
  ggplot(aes(x = forcats::fct_reorder(Chemical,
                                      relative_difference, .desc = FALSE))) +
  geom_linerange(aes(ymin = Css_httk_nr, ymax = Css_httk_r)) +
  labs(title = "Differences between restrictive & nonrestrictive clearance",
       y = "Css\n(mg/L)", x = "Chemicals") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(hjust = 0.5))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Css_difference_httk_lineplot.png"),
       bg = "white",
       height = 5, width = 7)

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Css_difference_httk_inkblot.png"),
       bg = "white",
       height = 5, width = 7)
###

>>>>

```{r library_loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ctxR)
library(ggvenn)
library(ggalluvial)
library(flextable)
library(reticulate)
library(patchwork)
source("data-raw/RC_comparison_bkgd.R") # httk & dawson_2021 loaded here
```

@@

```{r venn_diagrams}
member_list <- list(
  httk_3comp2 = unique(get_cheminfo(info = "DTXSID",
                                species = "human",
                                model = "3compartment2",
                                median.only = TRUE,
                                physchem.exclude = TRUE,
                                suppress.messages = TRUE)) %>% 
    suppressWarnings(),
  cvt = cvt %>% 
    filter(species %in% c("human", "rat"),
           conc_medium_normalized == "plasma") %>%
    pull(analyte_dtxsid)
)

@@

target_chems <- intersect(member_list$cvt, member_list$httk_3comp2)

@@

this_pk <- do_prefit(this_pk)
this_pk <- do_fit(this_pk, n_cores = 14)
```

```{r load-other}
# Here I want to load all other data and setup other data objects
# Kd/binding affinity data
load(paste0(Sys.getenv("OD_DIR"),
            "Profile/Documents/New-HTTK-raw-2019-05-06.RData")
)

# Chemical info and properties (averaging the latter by chemical and experimental/predicted)
target_chem_info <- get_chemical_details_batch(DTXSID = target_chems,
                                               rate_limit = 1L)
target_chem_properties <- get_chem_info_batch(DTXSID = target_chems, 
                                              rate_limit = 1L) %>%
  filter(name %in% c("LogKow: Octanol-Water",
                     "pKa Acidic Apparent",
                     "pKa Basic Apparent")) %>% 
  group_by(dtxsid, propType, propertyId) %>%
  summarize(Mean = mean(value, na.rm = TRUE),
            StDev = sd(value, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct()

# Bring together all chemical properties
target_chem_pw_predicted <- target_chem_properties %>%
  filter(propType == "predicted") %>%
  pivot_wider(names_from = propertyId,
              names_prefix = "pr_",
              id_cols = dtxsid,
              values_from = c("Mean", "StDev"))
target_chem_pw_experimental <- target_chem_properties %>%
  filter(propType == "experimental") %>%
  pivot_wider(names_from = propertyId,
              names_prefix = "ex_",
              values_from = c("Mean", "StDev"),
              id_cols = dtxsid)

target_chem_pw_full <- full_join(
  target_chem_pw_experimental,
  target_chem_pw_predicted
) %>%
  janitor::clean_names() %>%
  mutate(dtxsid,
         type_logkow = ifelse(is.na(mean_ex_logkow_octanol_water),
                              NA, "experimental"),
         logkow = ifelse(is.na(mean_ex_logkow_octanol_water),
                         mean_pr_logkow_octanol_water,
                         mean_ex_logkow_octanol_water),
         type_apKa = ifelse(is.na(mean_ex_pka_acidic_apparent),
                            NA, "experimental"),
         apKa = ifelse(is.na(mean_ex_pka_acidic_apparent),
                       mean_pr_pka_acidic_apparent,
                       mean_ex_pka_acidic_apparent),
         type_bpKa = ifelse(is.na(mean_ex_pka_basic_apparent),
                            NA, "experimental"),
         bpKa = ifelse(is.na(mean_ex_pka_basic_apparent),
                       mean_pr_pka_basic_apparent,
                       mean_ex_pka_basic_apparent),
         .keep = "unused") %>%
  mutate(type_logkow = ifelse(!is.na(logkow) & is.na(type_logkow),
                              "predicted",
                              ifelse(type_logkow == "experimental",
                                     "experimental", NA)),
         type_apKa = ifelse(!is.na(apKa) & is.na(type_apKa),
                            "predicted",
                            ifelse(type_apKa == "experimental",
                                     "experimental", NA)),
         type_bpKa = ifelse(!is.na(bpKa) & is.na(type_bpKa),
                            "predicted",
                            ifelse(type_bpKa == "experimental",
                                     "experimental", NA))) %>%
  mutate(logkow_sd = ifelse(type_logkow == "experimental",
                            st_dev_ex_logkow_octanol_water,
                            st_dev_pr_logkow_octanol_water),
         apKa_sd = ifelse(type_apKa == "experimental",
                          st_dev_ex_pka_acidic_apparent,
                          st_dev_pr_pka_acidic_apparent),
         bpKa_sd = ifelse(type_bpKa == "experimental",
                          st_dev_ex_pka_basic_apparent,
                          st_dev_pr_pka_basic_apparent)) %>%
  select(-starts_with("st_dev")) %>%
  distinct()
# Remove the parts of this final dataset
rm(target_chem_pw_experimental, target_chem_pw_predicted)

# Predictions from httk(code retained but not run after 2.20.2025)
restrictive_preds <- get_httk_preds(
  parameters = this_param_rest_l,
  pk_obj = this_pk,
  species = c("rat", "human")
) %>% rename_with(
  .fn = ~paste0(.x, "_restrictive"),
  .cols = starts_with("httk_")
)

names(restrictive_preds)

all_httk_preds <- inner_join(restrictive_preds, nonrestrictive_preds) %>%
  filter_targets()
rm(restrictive_preds, nonrestrictive_preds)
# 
readr::write_csv(all_httk_preds,
                 file = paste0(Sys.getenv("FIG_DIR"),
                               "RestrictiveClearance_Analysis/",
                               "CURRENT_all_httk_preds.csv"))

all_httk_preds <- readr::read_csv(file = paste0(Sys.getenv("FIG_DIR"), 
                                                "RestrictiveClearance_Analysis/",
                                                "CURRENT_all_httk_preds.csv"))

# Get the winning model (including flats) and only include chemicals in target_chems
this_winmodel <- get_winning_model(this_pk) %>%
    filter_targets()

# TKstats for 1-/2-compartment model
this_tkstats <- eval_tkstats(this_pk,
                             model = c("model_1comp", "model_2comp"),
                             finite_only = TRUE,
                             dose_norm = TRUE
) %>%
  filter_targets()
colnames(this_tkstats) <- gsub("/", "_", colnames(this_tkstats))
this_tkstats <- this_tkstats %>%
  inner_join(this_winmodel)

@@

```{r load_Kd}
new.raw.httk.data %>%
  distinct(DTXSID, Affinity.Kd.Med, Affinity.Kd.Low, Affinity.Kd.High,
           Fit) %>%
  na.omit() %>%
  mutate(Target_Chemical = (DTXSID %in% target_chems)) %>%
ggplot(
       aes(
         x = forcats::fct_reorder(DTXSID, Affinity.Kd.Low)
         )) +
  geom_linerange(aes(
    ymin  = Affinity.Kd.Low,
    ymax = Affinity.Kd.High,
    color = Target_Chemical
  ), alpha = 0.4) +
  geom_point(aes(
    y = Affinity.Kd.Med, color = Target_Chemical
  ), size = 0.75) +
  # facet_grid(cols = vars(Fit), space = "free_x", scales = "free_x") +
  scale_y_log10(labels = scales::label_log(digits = 1)) +
  scale_color_manual(values = c("black", "green4")) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.position = "bottom") +
  labs(title = "Kd estimates across chemicals", x = NULL)

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Kd_estimates_initialsummary.png"),
       bg = "white",
       height = 4, width = 9
)
```

@@

```{r lipinksi_test_data}

target_chem_properties %>%
  ggplot(aes(x = propType, y = Mean)) +
  geom_violin() +
  geom_line(aes(group = dtxsid), color = "grey10", alpha = 0.1) +
  geom_hline(yintercept = 5, linetype = "longdash") +
  facet_grid(scales = "free_x",
             rows = vars(propertyId)) +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(hjust = 1, angle = 45))

@@

target_chem_pw_full %>%
  ggplot(aes(x = forcats::fct_reorder(dtxsid, logkow))) +
  geom_pointrange(aes(y = logkow, ymin = logkow - logkow_sd,
                      ymax = logkow + logkow_sd), size = 0.1) +
  facet_grid(cols = vars(type_logkow)) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
target_chem_pw_full %>%
  ggplot(aes(x = logkow)) +
  geom_histogram(color = "grey50", fill = "grey5",
                 binwidth = 1) +
  theme_bw()

@@

```{python get_H_bonds}
import rdkit
from rdkit import Chem
from rdkit.Chem import Descriptors
from rdkit.Chem import rdMolDescriptors

@@

py_h_donors = [Chem.rdMolDescriptors.CalcNumLipinskiHBD(x) for x in mol_from_smiles]
py_h_acceptors = [Chem.rdMolDescriptors.CalcNumLipinskiHBA(x) for x in mol_from_smiles]
py_tpsa = [Chem.rdMolDescriptors.CalcTPSA(x, False, True) for x in mol_from_smiles]
py_rotbonds = [Chem.rdMolDescriptors.CalcNumRotatableBonds(x) for x in mol_from_smiles]
```

@@

target_chem_summary <- target_chem_info %>%
  distinct(dtxsid, monoisotopicMass, smiles) %>%
  right_join(data.frame(smiles = smiles_list)) %>%
  bind_cols(data.frame(H_donors = py_to_r(py$py_h_donors),
                       H_acceptors = py_to_r(py$py_h_acceptors),
                       TPSA = py_to_r(py$py_tpsa),
                       Rot_Bonds = py_to_r(py$py_rotbonds))
            ) %>%
  left_join(
    target_chem_pw_full
  ) %>%
  rowwise() %>%
  mutate(logD_7.4 = httk::calc_dow(dtxsid = dtxsid, pH = 7.4))

target_chem_lipinski <- target_chem_summary %>%
  select(-c(smiles)) %>%
  group_by(dtxsid) %>%
  mutate("Mass <= 500 Da" = (monoisotopicMass <= 500),
          "H-donors < 5" = (H_donors < 5),
          "H-acceptors < 10" = (H_acceptors < 10),
          "logP < 5" = (logkow < 5),
          .keep = "none") %>%
  ungroup() %>%
  mutate(Lipinski = rowSums(dplyr::select(., where(is.logical))) >= 3)

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Alluvium_Lipinski.png"),
       bg = "white",
       height = 5, width = 6
)

# Molecular descriptors in bins
bin_size <- 2
logP_plot <- target_chem_summary %>%
  distinct(dtxsid, logkow) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(logP_bin = factor(logkow %/% bin_size * bin_size)) %>%
  ggplot(aes(x = logP_bin, y = Funbound.plasma)) +
  labs(x = "logKow", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 50
TPSA_plot <- target_chem_summary %>%
  distinct(dtxsid, TPSA) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(TPSA_bin = factor(TPSA %/% bin_size * bin_size)) %>%
  ggplot(aes(x = TPSA_bin, y = Funbound.plasma)) +
  labs(x = "Topological Polar Surface Area", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 2
RotBond_plot <- target_chem_summary %>%
  distinct(dtxsid, Rot_Bonds) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(Rot_Bonds_bin = factor(Rot_Bonds %/% bin_size * bin_size)) %>%
  ggplot(aes(x = Rot_Bonds_bin, y = Funbound.plasma)) +
  labs(x = "Rotatable Bonds", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 150
MW_plot <- target_chem_summary %>%
  distinct(dtxsid, monoisotopicMass) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(MW_bin = factor(monoisotopicMass %/% bin_size * bin_size)) %>%
  ggplot(aes(x = MW_bin, y = Funbound.plasma)) +
  labs(x = "Molecular Mass", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 4
HBA_plot <- target_chem_summary %>%
  distinct(dtxsid, H_acceptors) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(HBA_bin = factor(H_acceptors %/% bin_size * bin_size)) %>%
  ggplot(aes(x = HBA_bin, y = Funbound.plasma)) +
  labs(x = "Hydrogen bond acceptors", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 2
HBD_plot <- target_chem_summary %>%
  distinct(dtxsid, H_donors) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(HBD_bin = factor(H_donors %/% bin_size * bin_size)) %>%
  ggplot(aes(x = HBD_bin, y = Funbound.plasma)) +
  labs(x = "Hydrogen bond donors", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

bin_size <- 200
target_chem_summary %>%
  distinct(dtxsid, logD_7.4) %>% 
  left_join(fup_df, by = join_by(dtxsid == Chemical)) %>%
  mutate(logD_bin =  factor(logD_7.4 %/% bin_size * bin_size)) %>%
  ggplot(aes(x = logD_bin, y = Funbound.plasma)) +
  labs(x = "logD_7.4", y = "fup") +
  ggbeeswarm::geom_beeswarm() +
  theme_bw()

(MW_plot + TPSA_plot + HBA_plot) / (logP_plot + RotBond_plot + HBD_plot)

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "PhysChemProperties_rdkit.png"),
       bg = "white",
       height = 6, width = 8
)

@@

```{r httk_pr_rmsle}

@@

combined_preds <- all_httk_preds %>%
  mutate(Reference = as.character(Reference)) %>%
  left_join(winning_preds) %>%
  filter(!is.na(model))

combined_RMSLE <- combined_preds %>%
  ungroup() %>%
  mutate(
    restrictive_SLE = (log1p(httk_Preds_restrictive) - log1p(Conc))^2,
    nonrestrictive_SLE = (log1p(httk_Preds_nonrestrictive) - log1p(Conc))^2,
    invivoPK_SLE =  (log1p(Conc_est) - log1p(Conc))^2 
  ) %>%
  group_by(Chemical, Species, Route) %>%
  summarize(RMSLE_restrictive = sqrt(mean(restrictive_SLE, na.rm = TRUE)),
            RMSLE_nonrestrictive = sqrt(mean(nonrestrictive_SLE, na.rm = TRUE)),
            RMSLE_invivoPKfit = sqrt(mean(invivoPK_SLE, na.rm = TRUE)))

@@

combined_RMSLE %>%
  ungroup() %>%
  mutate(comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
     between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive~ "More than"
  )) %>% 
  ggplot(aes(x = Chemical, y = RMSLE_invivoPKfit)) +
  geom_point(color = "green4") +
  geom_point(aes(y = RMSLE_nonrestrictive), shape = 6, size = 0.5) +
  geom_point(aes(y = RMSLE_restrictive), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = RMSLE_nonrestrictive, ymax = RMSLE_restrictive)) +
  facet_grid(~factor(comparison_group,
                     levels = c("Less than", "Between", "More than")),
             scales = "free_x", space = "free_x") +
  labs(y = "RMSLE") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank(),
        axis.title = element_text(size = 14),
        strip.text = element_text(size = 12))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "RMSLE_comparison_fullChart.png"),
       bg = "white",
       height = 3.8, width = 10
)# Number of chemicals we have RMSLE's for
combined_RMSLE %>% pull(Chemical) %>% unique() %>% length()

@@

```{r relative-likelihood}

# AIC and relative likelihood
this_rll_probs <- AIC(this_pk) %>%
  filter_targets() %>%
  distinct(Chemical, Species, model, AIC) %>%
  pivot_wider(names_from = model, values_from = AIC) %>%
  rowwise() %>% 
  mutate(minAIC = min(c_across(starts_with("model_")), na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(across(starts_with("model_"), \(x) {
    signif(exp((minAIC - x)/2) * 100, 4)
  })) %>%
  mutate(across(starts_with("model_"), \(x) {
    ifelse(x < 1E-6, 0, x)
  })) %>%
  inner_join(this_nonflat_winmodel %>% distinct(Chemical, Species, model)) %>%
  ungroup() %>%
  filter(model_1comp >= 99 | model_2comp >= 99,
        model_flat < 1, Chemical %in% target_chems)

@@

# LL from pre-calculated httk predictions
restrictive_ll <- all_httk_preds %>% 
  rename(Conc_est = httk_Preds_restrictive,
         Time_trans = Time) %>%
  select(!httk_Preds_nonrestrictive) %>%
  distinct() %>%
  group_by(Chemical, Species) %>%
  nest(.key = "observations") %>%
  inner_join(coef(this_pk) %>% semi_join(this_winmodel) %>% distinct()) %>%
  rowwise() %>%
  mutate(log_likelihood = log_likelihood(
    par = coefs_vector,
    data = observations,
    data_sigma_group = observations$data_sigma_group,
    dose_norm = FALSE,
    log10_trans = FALSE,
    negative = FALSE,
    includes_preds = TRUE,
    detect
  )) %>%
  select(!observations) %>%
  mutate(model = "httk_3comp2_restrictive",
         AIC = -2 * log_likelihood)

nonrestrictive_ll <- all_httk_preds %>% 
  rename(Conc_est = httk_Preds_nonrestrictive,
         Time_trans = Time) %>%
  select(!httk_Preds_restrictive) %>%
  distinct() %>%
  group_by(Chemical, Species) %>%
  nest(.key = "observations") %>%
  inner_join(coef(this_pk) %>% semi_join(this_winmodel) %>% distinct()) %>%
  rowwise() %>%
  mutate(log_likelihood = log_likelihood(
    par = coefs_vector,
    data = observations,
    data_sigma_group = observations$data_sigma_group,
    dose_norm = FALSE,
    log10_trans = FALSE,
    negative = FALSE,
    includes_preds = TRUE,
    detect
  )) %>%
  select(!observations) %>%
  mutate(model = "httk_3comp2_nonrestrictive",
         AIC = -2 * log_likelihood) # Simple formula when there are no optimized params

httk_AIC <- bind_rows(restrictive_ll, nonrestrictive_ll)

httk_AIC %>%
  distinct(Chemical, Species, model, AIC) %>%
  pivot_wider(names_from = model, values_from = AIC) %>%
  rowwise() %>% 
  mutate(minAIC = min(c_across(starts_with("httk_")), na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(across(starts_with("httk_"), \(x) {
    signif(exp((minAIC - x)/2) * 100, 4)
  })) %>%
  mutate(across(starts_with("httk_"), \(x) {
    ifelse(x < 1E-6, 0, x)
  })) %>% View()

@@

combined_model_weights %>%
  ggplot(aes(x = gamma_weight, color = rmsle_comparison_group)) +
  geom_freqpoly() +
  labs(x = "Inverse Confidence",
       color = "RMSLE relative to httk predictions") +
  facet_grid(rows = vars(rmsle_comparison_group)) +
  theme_bw()

@@

```{r combine_tkstats}
# How many overlapping chemicals have winning models that aren't flat?
count(
  this_winmodel,
  model
) # 153 = 97 1-compartment + 56 2-compartment

@@

this_css_df <- inner_join(
  css_httk_batch(this_param_rest_l, restrictive = TRUE),
  css_httk_batch(this_param_nrest_l, restrictive = FALSE)
)

# Table
this_css_df %>%
  count(Css_httk_r == Css_httk_nr,
        Css_httk_r - Css_httk_nr < 0.1 & Css_httk_r - Css_httk_nr > 0,
        Css_httk_r - Css_httk_nr >= 0.1) %>%
  rename("No difference" = "Css_httk_r == Css_httk_nr",
         "Less than 0.1" = "Css_httk_r - Css_httk_nr < 0.1 & Css_httk_r - Css_httk_nr > 0",
         "At least 0.1" = "Css_httk_r - Css_httk_nr >= 0.1")

# Chart with relative differences (think steps) of Css between clearance types
this_css_df %>%
  mutate(relative_difference = log1p(Css_httk_r) - log1p(Css_httk_nr)) %>%
  count(relative_difference == 0, Css_httk_r == 0) # 31 Chemicals have zero difference
###
this_css_df %>%
  mutate(relative_difference = log1p(Css_httk_r) - log1p(Css_httk_nr)) %>%
  filter(relative_difference > 0) %>%
  ggplot(aes(x = forcats::fct_reorder(Chemical,
                                      relative_difference, .desc = FALSE))) +
  geom_linerange(aes(ymin = Css_httk_nr, ymax = Css_httk_r),
                 linewidth = 1.25) +
  labs(title = "Differences between restrictive & nonrestrictive clearance",
       y = "Difference in Css\n(mg/L)", x = "Chemicals") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 18),
        axis.title = element_text(size = 14),
        axis.text.y = element_text(size = 10, color = "black"))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Css_difference_httk_lineplot.png"),
       bg = "white",
       height = 3.5, width = 11.5)

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Css_difference_httk_inkblot.png"),
       bg = "white",
       height = 5, width = 7)
###
```

The next step is to assess Css, halflife, and total clearance
deciding whether the invivoPKfit evidence points to more restrictive or non-restrictive
clearance. To do this we will create a dummy variable that will accumulate
the information.

  Restrictive  = -1  
  Nonrestrictive = +1  
  Undetermined = 0

We will keep track of this with labeled columns and sum at the end.

```{r full-eval}
this_evaluated_df <- this_css_df %>%
  inner_join(this_tkstats) %>% # Css
  inner_join(this_httk_clearance_df) %>% # CLtot and halflife
  inner_join(data.frame(Chemical = names(this_param_rest_l),
                        CLint = vapply(names(this_param_rest_l),
                                       \(x) {
                                         this_param_rest_l[[x]][["Clint"]]
                                       },
                                       FUN.VALUE = numeric(1L)
                        ))) %>%
  mutate(Css_clear = case_when(
    between(log2(Css.tkstats),
            log2(Css_httk_r)-1, log2(Css_httk_r)+1) &
      abs(log2(Css.tkstats/Css_httk_r)) < abs(log2(Css.tkstats/Css_httk_nr))~ -1,
    between(log2(Css.tkstats),
            log2(Css_httk_nr)-1, log2(Css_httk_nr)+1) ~ 1,
    .default = 0),
    halflife_clear = case_when(
      between(log2(halflife.tkstats),
              log2(restrictive_halflife)-1,
              log2(restrictive_halflife)+1) &
        abs(
          log2(halflife.tkstats/restrictive_halflife)
        ) < abs(
          log2(halflife.tkstats/nonrestrictive_halflife)
        ) | (nonrestrictive_halflife == restrictive_halflife) ~ -1,
      (log2(halflife.tkstats) < log2(nonrestrictive_halflife)+1) ~ 1,
      .default = 0
    ),
    CLtot_clear = case_when(
      between(log2(CLtot.tkstats),
              log2(restrictive_clearance)-1,
              log2(restrictive_clearance)+1) &
        abs(
          log2(CLtot.tkstats/restrictive_clearance)
        ) < abs(
          log2(CLtot.tkstats/nonrestrictive_clearance)
        ) ~ -1,
      between(log2(CLtot.tkstats),
              log2(nonrestrictive_clearance)-1,
              log2(nonrestrictive_clearance)+1) ~ 1,
      .default = 0
    )
  )

evaluated_df_summary <- this_evaluated_df %>%
  group_by(Chemical) %>%
  summarize(CLtot_clear = median(CLtot_clear),
            halflife_clear = median(halflife_clear),
            Css_clear = median(Css_clear)) %>%
  rowwise() %>%
  mutate(clearance_sum = sum(CLtot_clear, halflife_clear, Css_clear)) %>%
  ungroup()

count(evaluated_df_summary, clearance_sum < 0, clearance_sum > 0)
